\Large \bf Vigor: Toward Carrier-Grade Software Network Functions

### 2016-10-27 doc/paper/paper.tex
Fix typos.

To verify that this NF does not crash, one could use the approach in Dobrescu et al.~\cite{dobrescu2014software} and run exhaustive symbolic execution. The symbolic execution (SE) technique entails using a symbolic execution engine (SEE) to explore all paths through the code and check whether any failure conditions hold --- in our case crash conditions. The inputs to the program are marked ``symbolic,'' meaning that they subsume all possible input values, and then the SEE simulates the execution of the program while propagating the symbolic inputs along the dataflow paths into the program's variables. When facing a branch (e.g., an \code{if} or a \code{for} loop), the SEE clones the program state and takes each feasible path separately. Branches impose constraints on variable values (e.g., an \code{if} statement imposes restrictions on the values of the condition variable along the \code{else} branch); the SEE collects these constraints along each path. When evaluating a branch condition, the SEE asks an SMT constraint solver which of the branches are feasible, given the constraints collected so far on that path prefix.

One could use an SEE like KLEE~\cite{cadar2008klee} to symbolically execute the code above exhaustively --- once all paths are found to be crash-free, the crash-freedom proof is complete. To perform SE, it is necessary to provide symbolic models for the three network calls. Something simple suffices, such as \code{receive_packet} generating a new symbolic packet in \code{dst}; \code{can_send_packet} forking program state into two states, one for each possible return value; and \code{send_packet} doing nothing.

Alas, symbolic execution would never terminate. The reason is ``path explosion'': there is a large (often infinite) number of paths to explore, so the SEE cannot finish.  Path explosion results from cyclomatic complexity~\cite{mccabe:cyclomatic}: every branch condition doubles the number of paths, and every loop multiplies it by a factor; in our case, three \code{if} conditions inside an infinite loop give rise to an infinite number of paths.

{\bf Step 1: Add loop invariants.} Infinite (or very long) loops can be simplified using loop invariants~\cite[\S~2.1]{cormen2009introduction}. These are annotations that inform the SEE about invariant properties of the loop, such as ``no element of \code{packets} between \code{begin} and \code{end} has destination port equal to 9'', and are a good match for NF-style loops.  We added to KLEE support for havoc-ing~\cite{barnett2005boogie} automatically, to take advantage of such invariants and essentially reduce the infinite loop to a single abstract iteration that can be reasoned about in one pass.

@@

More formally:
\begin{enumerate}
    \item \(C\): A call sequence, or a run (\(C=Ci+Co\)) consists of a run input and run output. It summarizes a particular interaction scenario with the model. It consists of the calls to component API functions (with function names, arguments and return values), and leaves out all other details like other calls or unrelated path conditions.
    \item \(Ci\): The run input (\(Ci = <Fi,Fo>i=0..n-1 + Fin\)) is a set of: function names, function inputs(\(Fi\)) for all the functions in the sequence and function outputs(\(Fo\)) for all but the last call.
    \item \(Co\): The run output (\(Co = Fon\)) is the output of the last API function call in the sequence.
    \item \(Fi\): The function input (\(Fi = fun\_name + args + arg\_pointee\_vals_before + constraints\)) consists of the function arguments (possibly symbolic), and constraints bounding these arguments (specifically, the transitive closure of all the constraints involving the symbols mentioned in the formulas for the arguments). For a pointer argument (which must point to a concrete address --- practical limitation), the pointee must be considered according to the \S\ref{sec:handling-pointers}. We use the numerical value of the pointer to detect aliasing between the variables, and then throw it away, since we forbid pointer comparison anyway.
    \item \(Fo\): The function output (\(Fo = ret\_val + pointee\_vals\_after + constraints\)) is its return value and the pointee values for the output pointer arguments, as well as the constraints for all the symbols involved (transitive closure, same as above).
\end{enumerate}

@@

\paragraph{SE verifier} We built our verifier for speculative SE by modifying an open source C/C++ SEE Klee~\cite{cadar2008klee}. By default Klee terminates execution paths that are too long or take too much memory. We disabled this feature to ensure it explores all the paths. We also extended KLee with three new features:
Partial loop invariant support. The analysis finds automatically the variables that may change in the loop, and havocs them. Currently, a developer have to manually insert the assertions and assumptions according to the loop invariant definition.
Dynamic pointer access control. The patch provides primitive intrinsics that allow a component developer to open and close access to a pointer between the API calls.
API call sequence dump. This patch summarizes each control path in the program extracting the component API calls and dumps them for the Validator.

@@

\paragraph{Validator} We wrote this third tool completely from scratch. Validator independently analyzes each API calls sequence. We exploit the data parallelism using GNU parallel tool~\cite{Tange2011a}. For each API call sequence in the Klee dump, Validator first transforms it into VeriFast task. A VeriFast task is essentially a C source file, containing the sequence of calls. The Validator enriches the sequence with:
\begin{itemize}
    \item Symbolic variables used in the sequence.
    \item The constraints that form the sequence context.
    \item Lemmas to help formal verification.
\end{itemize}
Next, it runs VeriFast to check that the specification permits this particular sequence of calls under the provided constraints. Finally, the Validator imports the specification output state \(O_s\) and checks the inclusion of the specification state \(O_s\) to the model state \(O_m\) using Z3~\cite{de2008z3} SMT solver (see Section~\ref{sec:model-validation-criteria}).

>>>>

To verify that this NF does not crash, one could use the approach of Dobrescu et al.~\cite{dobrescu2014software} and run exhaustive symbolic execution. The symbolic execution (SE) technique entails using a symbolic execution engine (SEE) to explore all paths through the code and check whether any failure conditions hold --- in our case crash conditions. The inputs to the program are marked ``symbolic,'' meaning that they subsume all possible input values, and then the SEE simulates the execution of the program while propagating the symbolic inputs along the dataflow paths into the program's variables. When facing a branch (e.g., an \code{if} or a \code{for} loop), the SEE clones the program state and takes each feasible path separately. Branches impose constraints on variable values (e.g., an \code{if} statement imposes restrictions on the values of the condition variable along the \code{else} branch); the SEE collects these constraints along each path. When evaluating a branch condition, the SEE asks an SMT constraint solver which of the branches are feasible, given the constraints collected so far on that path prefix.

One could use an SEE like KLEE~\cite{cadar2008klee} to symbolically execute the
code above exhaustively --- once all paths are found to be crash-free, the
crash-freedom proof is complete. To perform SE, it is necessary to provide
symbolic models for the three network calls. Something simple suffices, such as
\code{receive_packet} generating a new symbolic packet in \code{dst};
\code{can_send_packet} forking program state into two states for busy and idle
interface correspondingly; and \code{send_packet} doing nothing.

Alas, symbolic execution would never terminate. The reason is ``path explosion'': there is a large (often infinite) number of paths to explore, so the SEE cannot finish.  Path explosion results from cyclomatic complexity~\cite{mccabe:cyclomatic}: every branch condition doubles the number of paths, and every loop multiplies it by a factor; in our case, the infinite loop gives rise to an infinite number of paths.

{\bf Step 1: Add loop invariants.} Infinite (or very long) loops can be
simplified using loop invariants~\cite[\S~2.1]{cormen2009introduction}. These
are annotations that inform the SEE about invariant properties of the loop, such
as ``no element of \code{packets} between \code{begin} and \code{begin +
  queue_len} has destination port equal to 9'', and are a good match for
NF-style loops.  We added to KLEE support for havoc-ing~\cite{barnett2005boogie}
automatically, to take advantage of such invariants and essentially reduce the
infinite loop to a single abstract iteration that can be reasoned about in one
pass.

@@

More formally:
\begin{enumerate}
    \item \(C\): A call sequence, or a run (\(C=Ci+Co\)) consists of a run input and run output. It summarizes a particular interaction scenario with the model. It consists of the calls to component API functions (with function names, arguments and return values), and leaves out all other details like other calls or unrelated path conditions.
    \item \(Ci\): The run input (\(Ci = <Fi,Fo>i=0..n-1 + Fin\)) is a set of: function names, function inputs(\(Fi\)) for all the functions in the sequence and function outputs(\(Fo\)) for all but the last call.
    \item \(Co\): The run output (\(Co = Fon\)) is the output of the last API function call in the sequence.
    \item \(Fi\): The function input (\(Fi = fun\_name + args + arg\_pointee\_vals\_before + constraints\)) consists of the function arguments (possibly symbolic), and constraints bounding these arguments (specifically, the transitive closure of all the constraints involving the symbols mentioned in the formulas for the arguments). For a pointer argument (which must point to a concrete address --- practical limitation), the pointee must be considered according to the \S\ref{sec:handling-pointers}. We use the numerical value of the pointer to detect aliasing between the variables, and then throw it away, since we forbid pointer comparison anyway.
    \item \(Fo\): The function output (\(Fo = ret\_val + pointee\_vals\_after + constraints\)) is its return value and the pointee values for the output pointer arguments, as well as the constraints for all the symbols involved (transitive closure, same as above).
\end{enumerate}

@@

\paragraph{SE verifier} We built our verifier for speculative SE by modifying an open source C/C++ SEE KLEE~\cite{cadar2008klee}. By default KLEE terminates execution paths that are too long or take too much memory. We disabled this feature to ensure it explores all the paths. We also extended KLEE with three new features:
Partial loop invariant support. The analysis finds automatically the variables that may change in the loop, and havocs them. Currently, a developer have to manually insert the assertions and assumptions according to the loop invariant definition.
Dynamic pointer access control. The patch provides primitive intrinsics that allow a component developer to open and close access to a pointer between the API calls.
API call sequence dump. This patch summarizes each control path in the program extracting the component API calls and dumps them for the Validator.

@@

\paragraph{Validator} We wrote this third tool completely from scratch. Validator independently analyzes each API calls sequence. We exploit the data parallelism using GNU parallel tool~\cite{Tange2011a}. For each API call sequence in the KLEE dump, Validator first transforms it into VeriFast task. A VeriFast task is essentially a C source file, containing the sequence of calls. The Validator enriches the sequence with:
\begin{itemize}
    \item Symbolic variables used in the sequence.
    \item The constraints that form the sequence context.
    \item Lemmas to help formal verification.
\end{itemize}
Next, it runs VeriFast to check that the specification permits this particular sequence of calls under the provided constraints. Finally, the Validator imports the specification output state \(O_s\) and checks the inclusion of the specification state \(O_s\) to the model state \(O_m\) using Z3~\cite{de2008z3} SMT solver (see Section~\ref{sec:model-validation-criteria}).
